Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-06-19 ThesisJP/Introduction/introduction.tex
Final changes.

\subsection{Paper\,I}
\textbf{An Optimized GPU Implementation of the MVDR Beamformer for Active Sonar Imaging}\\
J.\:I.\:Buskenes, \textbf{J.\:P.\:\AA{}sen}, C.-I.\:C.\:Nilsen and A.\:Austeng\\
{\it IEEE Transactions on Oceanic Engineering, accepted for publication.}\\\\
The first paper describes in detail how the Capon beamformer is mapped to the GPU architecture. Even though the paper is written within the field of active \nom{sonar}{SOund Navigation And Ranging (usually under water)} imaging, the presented implementation is applicable to a range of active imaging systems. A similar discussion for cardiac ultrasound imaging can be found in \textbf{Paper\,VI}.  Active sonar imaging typically differs from medical ultrasound imaging by a less strict real-time requirement and fewer array elements. This makes it somewhat easier to reach the goal of real-time processing. The transmit beams used for sonar imaging are also much wider than the once used for ultrasound imaging. This makes techniques like the BS-Capon beamformer less attractive for sonar imaging. However, as shown in this paper, low-complexity methods are not needed to reach real time Capon beamforming for sonar imaging.

@@

\subsection{Paper\,III}
\textbf{Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance}\\
\textbf{J.\:P.\:\AA{}sen}, A.\:Austeng and S.\:Holm\\
{\it IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, accepted for publication.}\\\\
In the third paper we study the shift-invariant property of an imaging system based on the Capon beamformer. As mentioned, shift-invariant imaging is essential if the method is ever to be used in practice. The paper is based on an observation of aliasing artifacts when the Capon beamformer was used to image bright point scatterers \textit{in vitro} and in simulations. The intensity of the point scatterers were observed to oscillate as either they or the probe moved. In earlier work on Capon beamforming for medical ultrasound imaging the real-time nature of ultrasound imaging has often been ignored, and a large degree of oversampling on transmit might have been applied without this being clearly stated. Oversampling on transmit will reduce the acquisition frame rate, and is therefore not an option if the imaging system is used to image rapidly varying objects. Smooth presentation of probe motion is equally important since real-time user interaction is one of the key selling points of ultrasound imaging.

@@

Achieving higher processing speed for an excising algorithm by implementing it on a new architecture is not academically relevant in its own. A good analysis has to be given of how the algorithm was modified for the architecture, or how to utilize the gained speed. These are points we have tried to answer in both \textbf{Paper\,I}, \textbf{II},  \textbf{IV}, and \textbf{V}.

@@

\section{Multimedia content}
In \textbf{Paper II} and \textbf{III} there will be references to several videos that can be accessed online\footnote{folk.uio.no/jpaasen/thesis}.

>>>>

\subsection{Paper\,I}
\textbf{An Optimized GPU Implementation of the MVDR Beamformer for Active Sonar Imaging}\\
J.\:I.\:Buskenes, \textbf{J.\:P.\:\AA{}sen}, C.-I.\:C.\:Nilsen and A.\:Austeng\\
{\it IEEE Transactions on Oceanic Engineering, Jul 2014.}\\\\
The first paper describes in detail how the Capon beamformer is mapped to the GPU architecture. Even though the paper is written within the field of active \nom{sonar}{SOund Navigation And Ranging (usually under water)} imaging, the presented implementation is applicable to a range of active imaging systems. A similar discussion for cardiac ultrasound imaging can be found in \textbf{Paper\,VI}.  Active sonar imaging typically differs from medical ultrasound imaging by a less strict real-time requirement and fewer array elements. This makes it somewhat easier to reach the goal of real-time processing. The transmit beams used for sonar imaging are also much wider than the once used for ultrasound imaging. This makes techniques like the BS-Capon beamformer less attractive for sonar imaging. However, as shown in this paper, low-complexity methods are not needed to reach real time Capon beamforming for sonar imaging.

@@

\subsection{Paper\,III}
\textbf{Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance}\\
\textbf{J.\:P.\:\AA{}sen}, A.\:Austeng and S.\:Holm\\
{\it IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control, , vol. 61, no. 7, pp. 1152-1160, Jul 2014.}\\\\
In the third paper we study the shift-invariant property of an imaging system based on the Capon beamformer. As mentioned, shift-invariant imaging is essential if the method is ever to be used in practice. The paper is based on an observation of aliasing artifacts when the Capon beamformer was used to image bright point scatterers \textit{in vitro} and in simulations. The intensity of the point scatterers were observed to oscillate as either they or the probe moved. In earlier work on Capon beamforming for medical ultrasound imaging the real-time nature of ultrasound imaging has often been ignored, and a large degree of oversampling on transmit might have been applied without this being clearly stated. Oversampling on transmit will reduce the acquisition frame rate, and is therefore not an option if the imaging system is used to image rapidly varying objects. Smooth presentation of probe motion is equally important since real-time user interaction is one of the key selling points of ultrasound imaging.

@@

Achieving higher processing speed for an existing algorithm by implementing it on a new architecture is not academically relevant in its own. A good analysis has to be given of how the algorithm was modified for the architecture, or how to utilize the gained speed. These are points we have tried to answer in both \textbf{Paper\,I}, \textbf{II},  \textbf{IV}, and \textbf{V}.

@@

\section{Multimedia content}
In \textbf{Paper II} and \textbf{III} there will be references to several videos that can be accessed online\footnote{folk.uio.no/jpaasen/thesis}. There is one folder for each paper having multimedia content.
